Forecasting the 2024 U.S. Presidential Election through Poll Aggregation and Adjustments for Poll Quality

### 2024-11-04 paper/paper.qmd
update model and paper

```{r}
#| label: fig-bayesian-model
#| fig-cap: "Posterior Predictive Check for the Bayesian Model Predicting pct"
#| echo: false
#| warning: false
baye_model_data = just_harris_data
baye_model_data$pct = as.factor(baye_model_data$pct)
baye_model_data$pollscore = as.factor(baye_model_data$pollscore)
baye_model_data$days_taken_from_election = as.factor(baye_model_data$days_taken_from_election)
baye_model_data$methodology = as.factor(baye_model_data$methodology)
baye_model_data$pct <- as.numeric(baye_model_data$pct)
# Define the Bayesian model with brms
# test and train dataset

formula <- pct ~ pollscore + days_taken_from_election + sample_size + (1 | methodology) + (1 | state)

priors = normal(0, 2.5, autoscale = TRUE)

bayesian_model_1 <- stan_glmer(
  formula = formula,
  data = just_harris_data,
  family = gaussian(),
  prior = priors,
  prior_intercept = priors,
  seed = 123,
  cores = 4,
  adapt_delta = 0.95
)
pp_check(bayesian_model_1)
# summary(bayesian_model_1)
```

@@

### Train Test Validation
we implemented a train-test split validation approach to evaluate the performance of a Bayesian hierarchical model predicting percentage support (pct) for Kamala Harris. The dataset was divided into training (70%) and test (30%) sets to fit and validate the model, respectively. We used the stan_glmer function to fit the model on the training data, with predictors including pollscore, days_taken_from_election, sample_size, and random intercepts for methodology and state. After training, we generated predictions on the test set and calculated an R squared value to quantify the model's predictive accuracy. 
```{r}
#| warning: false
#| echo: false
# Split data into training and test sets
set.seed(123)
train_indices <- sample(seq_len(nrow(just_harris_data)), size = 0.7 * nrow(just_harris_data))
train_data <- just_harris_data[train_indices, ]
test_data <- just_harris_data[-train_indices, ]

# Fit the model on the training data
bayesian_model_train <- stan_glmer(
  formula = formula,
  data = train_data,
  family = gaussian(),
  prior = priors,
  prior_intercept = priors,
  seed = 123,
  cores = 4,
  adapt_delta = 0.95
)

@@

# Result {#sec-result}
```{r}
#| label: fig-bayesian-model-CI
#| echo: false
#| warning: false
# Plot random effects
plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
```

@@

```{r}
#| echo: false
#| warning: false

predictions2 <- posterior_predict(bayesian_model_1, newdata = just_harris_data)

predicted_means <- colMeans(predictions2)
predicted_intervals <- apply(predictions2, 2, quantile, probs = c(0.025, 0.975))

# Combine the results with the actual data
result_summary <- data.frame(
  Actual_PCT = just_harris_data$pct,
  Predicted_PCT = predicted_means,
  Lower_CI = predicted_intervals[1, ],
  Upper_CI = predicted_intervals[2, ]
)

ggplot(result_summary, aes(x = Actual_PCT, y = Predicted_PCT)) +
  geom_point() +
  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(
    title = "U.S. Election Prediction Results",
    x = "Actual PCT",
    y = "Predicted PCT"
  )

print(head(result_summary))

>>>>

```{r}
#| label: fig-bayesian-model
#| fig-cap: "Posterior Predictive Check for the Bayesian Model Predicting pct"
#| echo: false
#| warning: false
bayesian_model_1 = readRDS(here("models/bayesian_model_1.rds"))

pp_check(bayesian_model_1)
```

@@

### Train Test Validation
we implemented a train-test split validation approach to evaluate the performance of a Bayesian hierarchical model predicting percentage support (pct) for Kamala Harris. The dataset was divided into training (70%) and test (30%) sets to fit and validate the model, respectively. We used the stan_glmer function to fit the model on the training data, with predictors including pollscore, days_taken_from_election, sample_size, and random intercepts for methodology and state. After training, we generated predictions on the test set and calculated an R squared value to quantify the model's predictive accuracy. 
```{r}
#| echo: false
#| warning: false
# Split data into training and test sets
set.seed(123)
# train_indices <- sample(seq_len(nrow(just_harris_data)), size = 0.7 * nrow(just_harris_data))
# train_data <- just_harris_data[train_indices, ]
# test_data <- just_harris_data[-train_indices, ]
train_data = read_parquet(here("data/02-analysis_data/train_data.parquet"))
test_data = read_parquet(here("data/02-analysis_data/test_data.parquet"))

# Fit the model on the training data
bayesian_model_train = readRDS(here("models/bayesian_model_train.rds"))
# bayesian_model_train <- stan_glmer(
#   formula = formula,
#   data = train_data,
#   family = gaussian(),
#   prior = priors,
#   prior_intercept = priors,
#   seed = 123,
#   cores = 4,
#   adapt_delta = 0.95
# )

@@

# Result {#sec-result}
```{r}
#| label: fig-bayesian-model-CI
#| fig-cap: "Estimated Posterior Distribution for the Intercept Parameter in the Bayesian Model"
#| echo: false
#| warning: false
# Plot random effects
plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
```

@@

```{r}
#| label: tbl-bayesian-model-result-summary
#| tbl-cap: "Bayesian Model Result Summary"
#| echo: false
#| warning: false

predictions2 = posterior_predict(bayesian_model_1, newdata = just_harris_data)

predicted_means = colMeans(predictions2)
predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))

# Combine the results with the actual data
result_summary = data.frame(
  Actual_PCT = just_harris_data$pct,
  Predicted_PCT = predicted_means,
  Lower_CI = predicted_intervals[1, ],
  Upper_CI = predicted_intervals[2, ]
)

kable(head(result_summary, 10))
```

```{r}
#| label: fig-bayesian-model-result-CI
#| fig-cap: "U.S. Election Prediction Results"
#| echo: false
#| warning: false
ggplot(result_summary, aes(x = Actual_PCT, y = Predicted_PCT)) +
  geom_point() +
  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(
    title = "U.S. Election Prediction Results",
    x = "Actual PCT",
    y = "Predicted PCT"
  )
```

@fig-bayesian-model-result-CI shows that he model's predictive performance for Harris's election data appears generally robust, demonstrating a strong alignment between actual and predicted vote percentages. The majority of the data points cluster around the line, indicating reliable predictions, some variance is evident, particularly in the lower and higher actual percentage ranges. This suggests that while the model performs well on average, it may have larger uncertainty in its predictions at the extremes. @tbl-bayesian-model-result-summary also highlights that most predicted values align closely with actual values, supporting the model’s robustness in capturing the overall trend of election outcomes.
